---
title: "Project1"
author: "likhitha"
date: "2022-10-23"
output:
  html_document: default
  word_document: default
---

# Import the required libraries

```{r}
library(ggplot2)
library(tidyverse)
library(nycflights13)
```

```{r}
glimpse(flights)
```

Create a new dataset named 'myflights' in order to filter the carrier as 'UA'.

```{r}
myflights <- flights %>% 
  filter(carrier == 'UA') %>% 
  mutate(
    late = dep_delay>0,
    very_late = dep_delay>30
  )
glimpse(myflights)
```
```{r}
glimpse(weather)
```
Appending both the data sets ‘myflights’ and ‘weather’ into a new data set named ‘data’ using inner join.

```{r}
data <- myflights %>% inner_join(weather, by = c('origin','year','month','day','hour','time_hour'))
glimpse(data)
```



# Departure Delay Vs Time of day

Creating a temporary dataset named 'data1' that stores average departure delay grouped by hour.

```{r}
data1 <- data %>% group_by(time_of_day = hour) %>% summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE))
data1
# Plotting a scatter plot between time and average departure delay.
ggplot(data1, aes(y = avg_dep_delay, x = time_of_day))+
  geom_point()
```


```{r}
observed <- mean(data$dep_delay[data$hour > 16 ], na.rm = TRUE) - mean(data$dep_delay[data$hour <= 16], na.rm = TRUE)
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
sample.size = nrow(data)

#group.1.size = the number of observations in the first group
day = data %>% filter(hour > 16)
group.1.size = nrow(day)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram(bins = 60, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat("The difference is significant.\nThe departure delays at evening times of a day are greater than departure delays at morning times of a day.")
```


#Departure Delay Vs Time of year

Create a temporary dataset 'data1' that contains average dep_delays group_by month.

```{r}
data1 <- data %>% group_by(month) %>% summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE))
data1
ggplot(data1, aes(x = month, y = avg_dep_delay))+
  geom_line() + scale_x_discrete(limits = c("1","2","3","4","5","6","7","8","9","10","11","12"))
```
## Summer Spring

```{r}
observed <-  mean(data$dep_delay[data$month >= 6 & data$month <= 8], na.rm = TRUE) - mean(data$dep_delay[data$month >=3 & data$month <=5], na.rm = TRUE) 
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
summer_spring = data %>% filter(month >=3 & month <=8)
sample.size = nrow(summer_spring)

#group.1.size = the number of observations in the first group
summer = data %>% filter(month >= 6 & month <= 8)
group.1.size = nrow(summer)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram(bins = 60, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat("The difference is significant. The departure delays in Summer are greater than Spring.")
```


## Summer Autumn

```{r}
observed <-  mean(data$dep_delay[data$month >= 6 & data$month <= 8], na.rm = TRUE) - mean(data$dep_delay[data$month >=9 & data$month <=11], na.rm = TRUE) 
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
summer_autumn = data %>% filter(month >= 6 & month <=11)
sample.size = nrow(summer_autumn)

#group.1.size = the number of observations in the first group
summer = data %>% filter(month >= 6 & month <= 8)
group.1.size = nrow(summer)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram(bins = 60, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat('The difference is significant. The departure delays in summer are greater than departure delays in autumn.')
```

## Summer Winter

```{r}
observed <-  mean(data$dep_delay[data$month >= 6 & data$month <= 8], na.rm = TRUE) - mean(data$dep_delay[data$month >11 | data$month <=2], na.rm = TRUE) 
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
summer_winter = data %>% filter((month >= 6 & month <=8) | (month >11 | month <=2))
sample.size = nrow(summer_winter)

#group.1.size = the number of observations in the first group
summer = data %>% filter(month >= 6 & month <= 8)
group.1.size = nrow(summer)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram(bins = 60, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat('The difference is significant. The departure delays in summer are greater than departure delays in winter.')
```


## Spring Autumn

```{r}
observed <- mean(data$dep_delay[data$month >=3 & data$month <=5], na.rm = TRUE) - mean(data$dep_delay[data$month >= 9 & data$month <= 11], na.rm = TRUE)
observed
# Spring - Autumn

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
spring_autumn = data %>% filter((month >=9 & month <=11) | (month >=3 & month<=5))
sample.size = nrow(spring_autumn)

#group.1.size = the number of observations in the first group
spring = data %>% filter(month >= 3 & month <= 5)
group.1.size = nrow(spring)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram(bins = 60, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat('The difference is significant. The departure delays in spring are greater than autumn.')
```


## Spring - Winter

```{r}
observed <- mean(data$dep_delay[data$month >=3 & data$month <=5], na.rm = TRUE) - mean(data$dep_delay[data$month >11 | data$month <= 2], na.rm = TRUE)
observed
# Spring - Winter

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
spring_winter = data %>% filter((month > 11 | month <= 2) | (month >=3 & month<=5))
sample.size = nrow(spring_winter)

#group.1.size = the number of observations in the first group
spring = data %>% filter(month >= 3 & month <= 5)
group.1.size = nrow(spring)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram(color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```

```{r}
cat('The difference is significant. The departure delays in spring are greater than winter.')
```

## Winter Autumn

```{r}
observed <- mean(data$dep_delay[data$month >11 | data$month <= 2], na.rm = TRUE) - mean(data$dep_delay[data$month >=9 & data$month <=11], na.rm = TRUE) 
observed
# Winter - Autumn

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
winter_autumn = data %>% filter((month > 11 | month <= 2) | (month >=9 & month<=11))
sample.size = nrow(winter_autumn)

#group.1.size = the number of observations in the first group
winter = data %>% filter(month > 11 | month <= 2)
group.1.size = nrow(winter)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram(color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```

```{r}
cat('The difference is significant. The departure delays in winter are greater than autumn.')
```


# Departure Delay Vs Temperature

Create a temporary dataset,'data1' that contains average departure delays grouped by temp.

```{r}
data1 <- data %>% mutate(temp = round(temp))
data1 <- data1 %>% group_by(temp) %>% summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE))
data1
ggplot(data1, aes(x = temp, y = avg_dep_delay))+
  geom_line()
```



```{r}
observed <- mean(data$dep_delay[data$temp > 50 ], na.rm = TRUE) - mean(data$dep_delay[data$temp <= 50], na.rm = TRUE)
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
sample.size = nrow(data)

#group.1.size = the number of observations in the first group
temp1 = data %>% filter(temp > 50)
group.1.size = nrow(temp1)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram( bins = 30, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat('The difference is significance. The departure delays with temp>50 are greater than temp<=50.')
```


# Departure Delay Vs Wind speed

Create a temporary dataset, data1 that contains the average departure delay group by wind_speed.

```{r}
data1 <- data %>% group_by(wind_speed = round(wind_speed)) %>% summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE))
data1
ggplot(data1, aes(x = wind_speed, y = avg_dep_delay))+
  geom_line()
```

```{r}
observed <- mean(data$dep_delay[data$wind_speed > 20 ], na.rm = TRUE) - mean(data$dep_delay[data$wind_speed <= 20], na.rm = TRUE)
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
sample.size = nrow(data)

#group.1.size = the number of observations in the first group
ws = data %>% filter(wind_speed > 20)
group.1.size = nrow(ws)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram( bins = 30, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat('There is no significant difference. The departure delay for wind speed > 20 is equal to wind speed <=20.')
```


## Precipitation

Create a temporary dataset, data1 that contains average departure delay group by precip.

```{r}
data1 <- data %>% group_by(precip) %>% summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE))
data1
ggplot(data1, aes(x = precip, y = avg_dep_delay))+
  geom_point()
```



```{r}
observed <- mean(data$dep_delay[data$precip > 0.6 ], na.rm = TRUE) - mean(data$dep_delay[data$precip <= 0.6], na.rm = TRUE)
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
sample.size = nrow(data)

#group.1.size = the number of observations in the first group
pre = data %>% filter(precip > 0.6)
group.1.size = nrow(pre)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram( bins = 30, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat("The difference is significant. The departure delays with precip > 0.6 are greater than departure delays with precip <= 0.6.")
```


## Visibility

Create a temporary dataset, data1 that contains the average departure delay group by visib.

```{r}
data1 <- data %>% group_by(visib) %>% summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE))
data1
ggplot(data1, aes(x = visib, y = avg_dep_delay))+
  geom_point()
```
```{r}
observed <- mean(data$dep_delay[data$visib <= 2.5 ], na.rm = TRUE) - mean(data$dep_delay[data$visib > 2.5], na.rm = TRUE)
observed

#N = number of simulations we will use
N <- 10^4-1

set.seed(2000)

#sample.size = the number of observations in our sample
sample.size = nrow(data)

#group.1.size = the number of observations in the first group
vis = data %>% filter(visib <= 2.5)
group.1.size = nrow(vis)

#create a blank vector to store the simulation results
result <- numeric(N)

#use a for loop to cycle through values of i ranging from 1 to N
for(i in 1:N)
{

  index = sample(sample.size, size=group.1.size, replace = FALSE)

  result[i] = mean(data$dep_delay[index], na.rm = TRUE) - mean(data$dep_delay[-index], na.rm = TRUE)
}

#plot a histogram of the simulated differences
#add a vertical line at the observed difference
ggplot(data=tibble(result), mapping = aes(x=result)) + 
geom_histogram( bins = 30, color = 'white') +
geom_vline(xintercept = observed, color = "red")


#Calculate the p-value
(sum(result >= observed) + 1) / (N + 1)
```
```{r}
cat("The difference is significant. The departure delays with visibility <= 2.5 are greater than visibility > 2.5")
```

